---
title: "Correlation"
format: revealjs
---

```{r, echo=F}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(jtools))
suppressPackageStartupMessages(library(readxl))
```


## Correlation vs t-test

In statistics, there are tests for __differences__; and tests for __association__

* t-test is for __difference__ between groups
  * are women better than men at volleyball?
* correlation is for __association__ between variables
  * is more practice associated with better performance

---

### t-test

Women are better than men at volleyball.

```{r}
volleyball = sleep %>% 
  mutate(
    `volleyball skill` = sqrt(extra^2),
    Gender = factor(group, levels=c(1,2), labels=c("Men", "Women"))
    )
ggplot(volleyball, aes(y=`volleyball skill`, x=Gender, fill=Gender))+
  stat_summary(fun.y=mean, position=position_dodge(width=0.95), geom="bar")+
  theme_apa()+
  #ylab("score")+
  theme(axis.title.y = element_text(angle = 0, hjust = 0.5, vjust=0.5))+
  scale_fill_discrete(guide = "none")
  
```

---

### correlation

More practice is generally associated with better performance (with exceptions)

```{r}
set.seed(1)
prac = read_excel("practice_vs_performance.xlsx") 
prac2=prac
prac2$practice=prac2$practice+runif(10,-1.5,1.5)
prac2$performance[1] = 7.5
prac2$practice[1] = 0.5
prac2$performance[8] = 0.5
prac = prac %>% bind_rows(prac2)
ggplot(prac, aes(x=practice, y=performance))+
  theme_apa()+
  geom_point()+
  geom_smooth(se=F, method = 'lm')+
  scale_x_continuous(breaks=0:10, limits=c(0,10)) +
  scale_y_continuous(breaks=0:10, limits=c(0,10)) +
  theme(axis.title.y = element_text(angle = 0, hjust = 0.5, vjust=0.5))
```

---




